### 2011-06-02 doc/manual.tex
String and branching functions, multiplicity functions return formalElements, start on work on manual, contents of software package in the paper, comma fix in bibliography.

\documentclass[preprint,12pt]{article}

\usepackage{graphicx}
\usepackage{hyperref}

\usepackage{amsmath}
\usepackage{amssymb}

\usepackage{color}
\usepackage{listings}
\usepackage{textcomp}
\definecolor{listinggray}{gray}{0.9}
\definecolor{lbcolor}{rgb}{0.9,0.9,0.9}
\lstset{
%	backgroundcolor=\color{lbcolor},
        language=Mathematica,
	tabsize=2,
	rulecolor=,
        basicstyle=\scriptsize,
        upquote=true,
%        aboveskip={1.5\baselineskip},
        columns=fixed,
        showstringspaces=false,
        extendedchars=true,
%        breaklines=true,
        prebreak = \raisebox{0ex}[0ex][0ex]{\ensuremath{\hookleftarrow}},
%        frame=single,
        showtabs=false,
        showspaces=false,
        showstringspaces=false,
        identifierstyle=\ttfamily,
        keywordstyle=\color[rgb]{0,0,1},
        commentstyle=\color[rgb]{0.133,0.545,0.133},
        stringstyle=\color[rgb]{0.627,0.126,0.941},
}

\newcommand{\go}{\stackrel{\circ }{\mathfrak{g}}}
\newcommand{\ao}{\stackrel{\circ }{\mathfrak{a}}}
\newcommand{\co}[1]{\stackrel{\circ }{#1}}
\newcommand{\pia}{\pi_{\mathfrak{a}}}
\newcommand{\piab}{\pi_{\mathfrak{a}_{\bot}}}
\newcommand{\gf}{\mathfrak{g}}
\newcommand{\af}{\mathfrak{a}}
\newcommand{\bff}{\mathfrak{b}}
\newcommand{\afb}{\mathfrak{a}_{\bot}}
\newcommand{\hf}{\mathfrak{h}}
\newcommand{\hfg}{\hf_{\gf}}
\newcommand{\hfb}{\mathfrak{h}_{\bot}}
\newcommand{\pf}{\mathfrak{p}}

\begin{document}

\title{{\bf Affine.m} reference manual}
\author{Anton Nazarov}

\begin{abstract}

\end{abstract}

\section{Introduction}
\label{sec:introduction-1}

\section{Download and installation}
\label{sec:downl-inst}

\subsection{Contents of software package}
\label{sec:cont-softw-pack}

\section{Software package}
\label{package}
The package can be freely downloaded from {\it github}. Latest release can be found here. To get the development code use the command
\begin{lstlisting}[language=bash]
 git clone git://github.com/naa/Affine.git
\end{lstlisting}

Contents of the package:
\begin{verbatim}
    Affine/                                root folder
      demo/                                  demonstrations
        demo.nb                                demo notebook
        paper.nb                               code from the paper
      doc/                                 documentation folder
        figures/                             figures in paper and manual
          timing.pdf                           diagram showing performance
          branching.pdf                        branching coefficients
        bibliography.bib                     bibliographic database
        manual.pdf                           manual
        manual.tex                           manual source
        paper.pdf                            paper
        paper.tex                            paper source
        TODO.org                             list of issues
      src/                                 source folder
        affine.m                             main software package
      tests/                               unit tests folder
        tests.m                              unit tests
      README.markdown                      short description of software
\end{verbatim}

\section{Usage}
\label{sec:usage}

\section{Function reference}
\label{sec:reference}
\begin{lstlisting}

Expect::"usage"="Expect[comment,value,expression] is used for unit tests. If expression != value it throws exception.";

standardBase::"usage"=
    "standardBase[coordinates] represents weight vector in standard (Bourbaki) coordinates.\
    E.g. simple roots for B2 are standardBase[1,-1] and standardBase[0,1]";

finiteWeight::"usage"=
    "finiteWeight[dimension_?NumberQ,coordinates_standardBase] represents\ 
    vector in weight space of finite-dimensional Lie algebra.\n 
    finiteWeight[dimension] returns dimension of the space, where weight vector is embedded\ 
    (i.e. for sl_n it is n+1).\n 
    finiteWeight[standardBase] returns standard base coordinates of weight of finite-dimensional Lie algebra";

dimension::"usage"=
    "dimension[wg_?weightQ] returns the dimension of the space, where finite part of weight lives";

finitePart::"usage"=
    "finitePart[wg_?weightQ] returns finite part of weight";

standardBase::"usage"=
    "standardBase[wg_?weightQ] returns coorinates of finite part of finite part of weight in standard (Bourbaki) basis";

makeFiniteWeight::"usage"=
    "makeFiniteWeight[{coordinates__?NumberQ}] creates finiteWeight with given coordinates in standard base";

Dot::"usage"=Dot::"usage" <> "\n It is defined for weights of finite and affine Lie algebras";

Plus::"usage"=Plus::"usage" <> "\n It is defined for weights of finite and affine Lie algebras";

Equal::"usage"=Equal::"usage" <> "\n It is defined for weights of finite and affine Lie algebras";

Times::"usage"=Times::"usage"<>"\n\n Mutliplication by numbers is defined for the elements of weight space of affine and finite-dimensional Lie algebras.\ 
    \n For example makeFiniteWeight[{1,2,3}]*2==makeFiniteWeight[{2,4,6}]]";

affineWeight::"usage"=
    "affineWeight[dimension_?NumberQ,fw_finiteWeigt,level_?NumberQ, grade_?NumberQ] represents\ 
    vector in weight space of affine Lie algebra.\n 
    affineWeight[dimension] returns dimension of the space of real roots, where finite weight vector is embedded\ 
    (i.e. for sl_n it is n+1).\n 
    affineWeight[finitePart] returns finite part of weight as finiteWeight structure\n
    affineWeight[level] returns level of affine weight\n
    affineWeight[grade] returns grade of affine weight";

level::"usage"=
    "level[w_affineWeight] or w[level] returns level of weight of affine Lie algebra";

grade::"usage"=
    "grade[w_affineWeight] or w[grade] returns grade of weight of affine Lie algebra";

makeAffineWeight::"usage"= 
    "makeAffineWeight[fw_finiteWeight,level_?NumberQ,grade_?NumberQ] creates affine weight with the given finite part fw, level and grade\n
    makeAffineWeight[{fw__?NumberQ},level_?NumberQ,grade_?NumberQ] creates affine weight with the given finite part fw, level and grade
    ";

makeHashtable::"usage"=
    "makeHashtable[keys_List,values_List] creates hashtable from the list keys and list of values";
keys::"usage"="keys[hashtable] gives all the keys in hashtable";
values::"usage"="values[hashtable] gives all the values in hashtable (in the same order as keys)";

finiteRootSystem::"usage"=
    "finiteRootSystem[rank_Integer,{roots_finiteWeight}] represents root system of finite-dimensional Lie algebra.\n
    finiteRootSystem[rank] returns rank of the root system\n
    finiteRootSystem[dimension] returns the dimension of the space where the roors are realized as the vectors\n
    finiteRootSystem[simpleRoots] returns unsorted list of simple roots in the root system.\n
    finiteRootSystem[simpleRoot][n_Integer] returns n'th simple root"

makeFiniteRootSystem::"usage"=
    "makeFiniteRootSystem[{roots__finiteWeight}] creates finiteRootSystem structure from the List of simple roots\n
    makeFiniteRootSystem[{roots__finiteWeight}] creates  finiteRootSystem structure from the List of simple roots\n
    given as the lists of coordinates";

rank::"usage"=
    "rank[rs_?rootSystemQ] or rs[rank] returns rank of root system. Rank of affine root system is equal to rank of underlying \n
    finite-dimensional root system";

simpleRoot::"usage"=
    "simpleRoot[rs_?rootSystemQ][i_Integer] or rs[simpleRoot][i_Integer] returns simple root with the index i. For affine root system i can be zero";

simpleRoots::"usage"=
    "simpleRoots[rs_?rootSystemQ] or rs[simpleRoots] returns unsorted list of simple roots";

prependZeros::"usage"=
    "prependZeros[num_Integer,vec_finiteWeight] embeds vec to bigger space by prepending num zeros to its coordinates";

appendZeros::"usage"=
    "appendZeros[num_Integer,vec_finiteWeight] embeds vec to bigger space by appending num zeros to its coordinates";

CirclePlus::"usage"=CirclePlus::"usage" <> "\n Direct sum of finite-dimensional and affine Lie algebras can be specified as sum of root systems";

makeSimpleRootSystem::"usage"=
    "makeSimpleRootSystem[series, rank] creates root system of the algebra from given series with given rank.\n
    Exceptional Lie algebras are not supported yet";

Subscript::"usage"=Subscript::"usage" <>
    "\n Subscript[A|B|C|D|E,n_Integer] creates root system of the algebra from given series with given rank. \n
    It is an equivalent to makeSimpleRootsSystem, but looks better in notebook interface."

A::"usage"="represents root systems of series A";
B::"usage"="represents root systems of series B";
C::"usage"="represents root systems of series C";
D::"usage"="represents root systems of series D";
E::"usage"="represents root systems of series E";

weightQ::"usage"=
    "Weight predicate"

checkGrade::"usage"=
    "computations for affine Lie algebras are limited by some grade, by default it is equal to 10. To change it set rs[gradeLimit].\n
    checkGrade[rs_?rootSystemQ][w_?weightQ] checks if given weight has absolute value of grade less then gradeLimit. Returns true 
    for finite-dimensional case";

gradeLimit::"usage"=
    "Limits computation for affine Lie algebras, i.e. b2a=OverHat[Subscript[B,2]]; b2a[gradeLimit]=15";

reflection::"usage"=
    "reflection[x_finiteWeight] represents reflection in the hyperplane, orthogonal to the given root x\n
    reflection[x_affineWeight] represents affine reflection in the hyperplane, orthogonal to the given root x\n
    reflection[x][y] reflects y in the hyperplane, orthogonal to x"

coroot::"usage"="returns coroot of given root"

cartanMatrix::"usage"=
    "cartanMatrix[rs_?rootSystemQ] returns Cartan matrix of root system rs";

rootSystemQ::"usage"=
    "Predicate for root system data structures";

weylGroupElement::"usage"=
    "weylGroupElement[x__Integer][rs_?rootSystemQ] represents element of Weyl group composed of basic reflecetions x \n
    of root system rs. Example: weylGroupElement[1,2,1][makeSimpleRootSystem[B,2]] = s1 s2 s1";

fundamentalWeights::"usage"=
    "fundamentalWeights[rs_?rootSystemQ] calculates fundamental weights of given root system rs";

rho::"usage"=
    "rho[rs_?rootSystemQ] - Weyl vector of root system rs (sum of fundamental weights)";

toFundamentalChamber::"usage"=
    "toFundamentalChamber[rs_?rootSystemQ][vec_?weightQ] acts on a weight vec by simple reflections of rs till it gets to main Weyl chamber";

mainChamberQ::"usage"=
    "mainChamberQ[rs_?rootSystemQ][wg_?weightQ] tells if weights wg lies inside the main Weyl chamber of root system rs or not";

partialOrbit::"usage"=
    "partialOrbit[rs_finiteRootSystem][{weights__finiteWeight}] constructs Weyl partial orbit of given set of weights. \n
    It starts with the list of weights and reflects them away from main Weyl chamber.\n
    For w in fundamental chamber it gives the whole orbit.";

orbit::"usage"=
    "orbit[rs_?rootSystemQ][wg_?weightQ] returns the weight of Weyl orbit, containing given weight wg. \n
    Orbit is given as the list of lists starting with the weight in dominant Weyl chamber\n
    orbit[rs_?rootSystemQ][{wg_?weightQ}] works for a list of weights";

positiveRoots::"usage"=
    "positiveRoots[rs_?rootSystemQ] returns positive roots of root system rs";

dimension::"usage"=
    "dimension[rs_?rootSystemQ][hweight_?weightQ] returns dimension of Lie algebra highest weight representation";

weightSystem::"usage"=
    "weightSystem[rs_?rootSystemQ][highestWeight_?weightQ] returns the set of dominant weights in the highest weight module. \n
    The list is split in pieces by number of root substractions";

freudenthalMultiplicities::"usage"=
    "freudenthalMultiplicities[rs_finiteRootSystem][highestWeight_finiteWeight] returns hashtable with the multiplicities of 
    weights in the highest weight module";

orbitWithEps::"usage"="returns orbit with the determinants of Weyl reflections";

racahMultiplicities::"usage"=
    "racahMultiplicities[rs_?rootSystemQ] returns hashtable with the multiplicities of 
    weights in the highest weight module. It uses recurrent relation similar to Racah formula";

highestRoot::"usage"="returns highest root of root system";

makeAffineExtension::"usage"=
    "makeAffineExtension[fs_finiteRootSystem] creates root system of affine Lie algebra which 
    is extension of given finite-dimensional root system";

OverHat::"usage"=
    "OverHat[rs_finiteRootSystem] creates affine extension of root system rs, it is equivalent makeAffineExtension[rs], but uses mathematical notation in notebook interface";

affineRootSystem::"usage"=
    "affineRootSystem[rank_Integer,finiteRootSystem_finiteRootSystem,imaginaryRoot_affineWeight,realRoots_List] 
    represents root system of affine Lie algebra.\n
    affineRootSystem[rank] returns rank of the root system\n
    affineRootSystem[imaginaryRoot] returns imaginary (zeroth) root\n
    affineRootSystem[realRoots] returns list of real roots (roots of finite-dimensional subalgebra)
    affineRootSystem[dimension] returns the dimension of the space where the roots are realized as the vectors\n
    affineRootSystem[simpleRoots] returns unsorted list of simple roots in the root system.\n
    affineRootSystem[simpleRoot][n_Integer] returns n'th simple root"

zeroWeight::"usage"="zeroWeight[rs_?rootSystemQ] returns zero root of root system rs (zero of rs[dimension]-dimemsional space)";

marks::"usage"="marks[rs_affineRootSystem] returns marks of affine Lie algebra";

marks::"usage"="comarks[rs_affineRootSystem] returns comarks of affine Lie algebra";

weight::"usage"=
    "weight[rs_?rootSystemQ][labels__Integer] constructs weight defined by Dynkin labels";

dynkinLabels::"usage"=
    "dynkinLabels[rs_?rootSystemQ][wg_?weightQ] returns Dynkin labels of weight (coefficients of its decomposition to the sum of fundamental weights) ";

orthogonalSubsystem::"usage"=
    "orthogonalSubsystem[rs_?rootSystemQ,subs_?rootSystemQ] returns subset of positive roots of root system rs, which 
    are orthogonal to simple roots of subsystem subs";

projection::"usage"=
    "projection[rs_?rootSystemQ][weights__?weightQ] projects given weight to the root (sub)system\n
    projection[rs_?rootSystemQ][fe_formalElement] projects the formal elements";

formalElement::"usage"=
    "Datastructure to represent formal elements of the ring of characters (linear combinations of formal exponents of weights).\n
    Internally the data is held in hashtable.\n
    fe_formalElement[weight_?weightQ] returns multiplicity of a given weight (coefficient in front of Exp[weight])\n
    fe_formalElement[weights] returns list of weights\n
    fe_formalElement[multiplicities] returns list of multiplicities
    formalElements can be added, multiplied by number, Exp[wg_?weightQ] and by formaElements\n
    fe_formalElement[hashtable] returns formalElement's data as hashtable\n
    "

hashtable::"usage"=
    "hashtable[fe_formalElement] or fe[hashtable] returns weights and multiplicities of formal element as a hash table";

makeFormalElement::"usage"=
    "makeFormalElement[{weights___?weightQ},{multiplicities___?NumberQ}] creates formal element with given weights and corresponding multiplicities\n
    makeFormalElement[{weights__?weightQ}] creates formal element where weight multiplicity is calculated as the number of appearances of weight in arguments list"

subElement::"usage"=
    "subElement[fe_formalElement,{weights___?weightQ}] creates formalElement with the given subset of weights";

regularSubalgebra::"usage"=
    "regularSubalgebra[rs_finiteRootSystem][rootIndices__?NumberQ] returns root system of regular subalgebra obtained by removing all unlisted nodes \n
    from Dynkn diagram of the algebra";

simpleBranching::"usage"=
    "Calculate branching coefficients with simple algorithm, which constructs all the modules of subalgebra with Freudenthal formula";

anomalousWeights::"usage"="
    anomalousWeights[rs_?rootSystemQ][hweight_?weightQ] returns the formal element, consisting of anomalous weights";

fan::"usage"=
    "Constructs fan of the embedding";

ourBranching::"usage"=
    "Calculate branching coefficients with alternative algorithm";

branching2::"usage"=
    "Calculate branching coefficients with alternative algorithm without folding to main Weyl chamber";

drawPlaneProjection::"usage"=
    "drawPlaneProjection[axe1_,axe2_,f_formalElement] draws projection to the plane of axe1, axe2 (e.g. 1,2 or 3,4, or 1,4 etc) of some formal element
    \n(branching, multiplicities, tensor product coefficients)";

draw3dProjection::"usage"=
    "draw3dProjection[axe1_,axe2_,axe3_,f_formalElement] draws projection to the 3d space of axe1, axe2, axe3 (e.g. 1,2,3 or 1,3,4 etc) of some formal element
    \n(branching, multiplicities, tensor product coefficients)"

tensorProduct::"usage"=
    "calculate tensor product decomposition coefficients"

stringFunction::"usage"=
    "stringFunciton[rs_affineRootSystem][hweigth_affineWeight,wg_affineWeight] returns power series decomposition of string function"

branchingFunction::"usage"=
    "branchingFunciton[rs_affineRootSystem,subs_affineRootSystem][hweigth_affineWeight,wg_affineWeight] returns power series decomposition of branching function"

\end{lstlisting}

  \end{document}